---
title: "Occasions of lice treatments at aquaculture farms"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
date: "`r Sys.Date()`"
output:
  html_document:
    css: '~/github/coastrisk/src/style_coast.css'
    toc: true
    toc_float: true
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: 
  pdf_document:
    toc: true
---

```{r setup, include=FALSE, }
knitr::opts_chunk$set(echo = TRUE, message=FALSE)
source('~/github/coastrisk/src/common.r')
```

```{r additional libraries, inlcude = FALSE, message=FALSE}
library(data.table)
library(vroom)
library(rebus)
library(here)
```

# Load the data on lice treatment

The data was downloaded from the BarentsRisk website on 03.11.2021 from
https://www.barentswatch.no/nedlasting/fishhealth/treatments?lang=no



```{r}
lustreat <- read.csv(here("prep", "aquaculture", "data", "raw", "tiltak_mot_lakselus.csv"))
```

# Cleaning the lice treatment data

Note that i keep only the rows with chemical treatment (medikamentell tiltak). Chemical treatment has two types - bath and feed.
I remove 13 observations that do not have information on the municipality

```{r}
lustreat_prep <- lustreat |> 
  filter(Tiltak == "medikamentell") |> 
  select(-c(ArtsId, Rensefisk, Antall, Tiltak, Fylkesnummer, Fylke)) |> 
  janitor::clean_names() |> 
  rename(year = ar,
         week = uke,
         location_number = lokalitetsnummer,
         location_name = lokalitetsnavn,
         treatment_type = type_behandling,
         chemical = virkestoff,
         treatment_area = omfang,
         municip_number = kommunenummer,
         municip = kommune,
         prod_area_id = produksjonsomrade_id,
         prod_area_name = produksjonsomrade) |> 
  filter(!is.na(municip_number))
```


I need to visualize how many bath treatments were there per year and production area, and how many feed treatments.

# Clean the production areas names

I need to make them similar to how they are written in the other datasets.
For that I will import momb table (can be any other completed dataset) and take the production area identifiers from there.

```{r}
prod_areas <- read.csv("./data/output/momb_per_prodarea.csv")
#unique(prod_areas$prod_area_name)

prod_areas_df <- prod_areas |> 
  select(c(prod_area_id, prod_area_name)) |> 
  group_by(prod_area_id, prod_area_name) |> 
  slice_head()

```
Merge lice treatment table with prod_areas_df table:


```{r}
lustreat_prep2 <- lustreat_prep |> 
  select(-prod_area_name) |> 
  left_join(prod_areas_df, by = "prod_area_id")
```

Prepare the final data set with data on the number of bath treatments and feed treatments per production area and year (i take a sum of all events over a year)
Number of treatments (either bath or feed) per production area and year varies from 1 to 417

```{r}
treat_summary <- mosaic::fav_stats(lustreat_prep3$sum_treaments)
treat_summary
```

```{r}
lustreat_prep3 <- lustreat_prep2 |> 
  mutate(treatment_type = replace(treatment_type,
                                  treatment_type == "fôrbehandling",
                                  "Feed")) |> 
  mutate(treatment_type = replace(treatment_type,
                                  treatment_type == "badebehandling",
                                  "Bath")) |> 
  group_by(year,treatment_type, prod_area_id, prod_area_name) |> 
  summarize(sum_treaments = n())
  
```

How many years with data are there per production area? There somtimes 10 years with data per production area but also 3 to 9 years, meaning that lice treatment does not occur every year (or is not reported every year)

```{r}
obs_per_area <- lustreat_prep3 |> 
  group_by(prod_area_name, treatment_type) |> tally()

datatable(obs_per_area)
```


# Save the final table
```{r}
write_csv(lustreat_prep3, "./data/output/lice_treatments_per_prodarea.csv")
```

