---
title: "MOMB investigations per aquaculture location"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
date: "`r Sys.Date()`"
output:
  html_document:
    css: '~/github/coastrisk/src/style_coast.css'
    toc: true
    toc_float: true
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: 
  pdf_document:
    toc: true
---

```{r setup, include=FALSE, }
knitr::opts_chunk$set(echo = TRUE, message=FALSE)
source('~/github/coastrisk/src/common.r')
```

```{r additional libraries, inlcude = FALSE}
library(data.table)
library(vroom)
library(rebus)
library(here)
```


#############################
# Load mom-b examination data and municipalities' names


Load the momb data for whole Norway (sent by Fiskeidr in September 2021)
```{r}
momb <-read_excel("./data/raw/B-undersøkelser helelandet og historiske.xlsx",
                  sheet = 3)
```



# Load and prepare production areas data

I edit production areas table in the same way as for the other aquaculture indices, so we have similar datasets per index
```{r}
mcp_prod <- read.csv(file.path("./data/raw/municipality_productionarea.csv"),sep = " ")
```

```{r}
mcp_prod_prep <- mcp_prod %>% 
  rename(county_number = fylkenummer,
         municip_number = kommunenummer,
         municip_name = komname,
         total_mcp_area = Totalt_areal_inkl_12nm,
         sea_mcp_area = Havflate_12nm,
         prod_area_id = id,
         prod_area_name = name
         ) %>% 
  select(-knum) %>% 
  mutate(
    prod_area_name = str_replace_all(
      prod_area_name, pattern = "->", replacement = "-"
    )
  ) %>% 
  mutate(
    prod_area_name = str_replace_all(
      prod_area_name, pattern = fixed("+"), replacement = "-"
    )
  )

mcp_prod_prep2 <- mcp_prod_prep %>% 
  mutate(municip_number =
           ifelse(
             str_count(municip_number, pattern = regex("[:digit:]")) == 3,
                       str_c("0",municip_number),
                       municip_number
             )
         )
```


# Further preparation of momb data


```{r}
momb_prep <-  momb %>% 
  clean_names(.) |> 
  mutate(Year = lubridate::year(dato_provetaking)) %>% 
  select(-c("mom_type", "reference")) %>% 
  rename("kommune" = "lokalitets_kommune") 
  
```




I will try to import a data on municipalities' numbers and add the numbers to momb table
```{r}
municips <- read_excel("../administrative/fylker_kommuner_2019_2020_changes.xlsx")

municips_prep <-municips |> 
  select(-starts_with("Fylkes")) |> 
  clean_names() |> 
  mutate(kommunenr_2019 = ifelse(
  str_detect(kommunenr_2019, pattern = "^[[:digit:]]{3}$"), 
  str_c(kommunenr_2019, "0"),
  kommunenr_2019)
  )
```


How many rows do not have municipality information in the original momb data? We will have to just remove these rows.

```{r}
sum(is.na(momb_prep$kommune))
```


I have to rename some of the municipalities so that they have matches in municipalities' numbers table
```{r}
momb_prep3 <- momb_prep |> 
  filter(!is.na(kommune)) |> 
  mutate(kommune = replace(kommune, 
                           kommune == "BØ I NORDLAND",
                           "BØ")) |> 
    mutate(kommune = replace(kommune, 
                           kommune == "GÁIVUOTNA - KÅFJORD - KAIVUONO",
                           "KÅFJORD")) |> 
    mutate(kommune = replace(kommune, 
                           kommune == "BØ I NORDLAND",
                           "BØ"))|> 
  mutate(kommune = replace(kommune, 
                           kommune == "LOABÁK - LAVANGEN",
                           "LAVANGEN")) |> 
  mutate(kommune = replace(kommune, 
                           kommune == "UNJÁRGA - NESSEBY",
                           "NESSEBY")) |> 
  left_join(municips_prep[,c(1,2)],
            by = c("kommune" = "kommunenavn_2019")) |> 
  mutate(kommunenr_2019 = replace(kommunenr_2019, 
                           kommune == "HERØY I NORDLAND",
                           1818))  |> 
  mutate(kommunenr_2019 = replace(kommunenr_2019, 
                           kommune == "HERØY I MØRE OG ROMSDAL",
                           1515)) 
  
```

Remove duplicates, because some of the 2019 municipalities were repeated 2-3 times, those that have merged into one municipality in 2020

```{r}
momb_prep4 <- momb_prep3 |> 
  group_by(loknr,dato_provetaking,kommune, Year, kommunenr_2019) |> 
  slice_head()

```

Merge momb table with production areas table, using municipalities' numbers 
```{r}
momb_prep5 <- momb_prep4 |> 
  mutate(municip_name = str_to_sentence(kommune)) |> 
  rename(year = Year) |> 
  select(-kommune) |> 
  mutate(kommunenr_2019 = as.character(kommunenr_2019)) |> 
  left_join(mcp_prod_prep2[,c(2,4:7)], by = c("kommunenr_2019" = "municip_number"))
  
```

Mismatches again are:
```{r}
mismatch2 <- filter(momb_prep5, is.na(prod_area_name)) |> group_by(municip_name, kommunenr_2019) |> tally()


mismatch2

```
Only 11 mismatches

The problem with mismatches is that momb data used newest municipality names (after 2020 changes), while production areas table uses per-2020 names. Some municipalities, like ALVER formed after merging several municipalities - Lindas, Radoy, Melas. Which one is meant in the momb table is not possible to find out, so I will just assume it is one of the free (e.g., Lindas). 

Here is the summary how i will replace the names of municipalities:

```{r}
replace_municips <- tibble(
  new_names = mismatch2$municip_name,
  old_names = c("Lindås",
                "Fusa",
                "Norddal",
                 "Halsa",
  "Eide",
  "Flora",
  "Lillesand",
  "Vikna",
  "Orkdal",
  "Berg",
  "Eid"),
  
  old_number = c(1263,
                 1241,
                 1524,
                 1571,
                 1551,
                 1401,
                 0926,
                 5050,
                 5024,
                 1929,
                 1443)
)

replace_municips
```

I go back to table momb_prep4 and change the municipalities numbers for these 11 municipalities, such that we get match with production areas
```{r}
momb_prep6 <- momb_prep4 |> 
  mutate(municip_name = str_to_sentence(kommune))  |> 
  mutate(kommunenr_2019 = replace(kommunenr_2019,
                                  municip_name == "Alver",
                                  1263)) |> 
  mutate(kommunenr_2019 = replace(kommunenr_2019,
                                  municip_name == "Bjørnafjorden",
                                  1241)) |> 
  mutate(kommunenr_2019 = replace(kommunenr_2019,
                                  municip_name == "Fjord",
                                  1524)) |> 
  mutate(kommunenr_2019 = replace(kommunenr_2019,
                                  municip_name == "Heim",
                                  1571)) |> 
  mutate(kommunenr_2019 = replace(kommunenr_2019,
                                  municip_name == "Hustadvika",
                                  1551)) |> 
  mutate(kommunenr_2019 = replace(kommunenr_2019,
                                  municip_name == "Kinn",
                                  1401)) |> 
  mutate(kommunenr_2019 = replace(kommunenr_2019,
                                  municip_name == "Lillesand",
                                  paste0(0, 926))) |> 
  mutate(kommunenr_2019 = replace(kommunenr_2019,
                                  municip_name == "Nærøysund",
                                5050)) |> 
  mutate(kommunenr_2019 = replace(kommunenr_2019,
                                  municip_name == "Orkland",
                                  5024)) |> 
  mutate(kommunenr_2019 = replace(kommunenr_2019,
                                  municip_name == "Senja",
                                  1929)) |> 
  mutate(kommunenr_2019 = replace(kommunenr_2019,
                                  municip_name == "Stad",
                                  1443)) |> 
  rename(year = Year) |> 
  select(-kommune) |> 
  mutate(kommunenr_2019 = as.character(kommunenr_2019)) |> 
  left_join(mcp_prod_prep2[,c(2,4:7)], by = c("kommunenr_2019" = "municip_number"))
  
```

Check if there are still mismatches:

```{r}
mismatch3 <- filter(momb_prep6, is.na(prod_area_name)) |> group_by(municip_name, kommunenr_2019) |> tally()

mismatch3

```
Final table 
```{r}
momb_prep7 <- momb_prep6 |> 
  rename(municip = kommune,
         location_number = loknr,
         location_name = lok_navn,
         momb_status = miljo_tilstand,
         date = dato_provetaking,
         municip_number_2019 = kommunenr_2019)
```


Save the final table
```{r}
write.csv(momb_prep7, "./data/output/momb_per_prodarea.csv", row.names = F)
```



###################################################################
Here is just some other info on the old and new names of the municipalities

```{r}
old_names <- c("Lindås",
               "Fusa",
               "Bø",
               "c(\Fauske\, \Fuossko\)",
               "Norddal",
               "c(\Gáivuotna\, \Kaivuono\, \Kåfjord\)",
               "c(\Hamarøy\, \Hábmer\)",
               "c(\Harstad\, \Hárstták\)",
               "Hemne",
               "Herøy i møre og romsdal",
               "Herøy i nordland",
               "Eide",
               "Flora",
               "c(\Lavangen\, \Loabák\)",
               "Vikna",
               "c(\Nordreisa\, \Raisi\, \Ráisa\)",
               "Orkdal",
               "Lenvik",
               "Sør-Varanger",
               "c(\Sortland\, \Suortá\)",
               "Eid",
               "c(\Nesseby\, \Unjárga\)")
```

```{r}
update_names <- tibble(
  new_name = mismatch$municip_name,
  old_name = c("Lindås",
               "Fusa",
               "Bø",
               "c(\Fauske\, \Fuossko\)",
               "Norddal",
               "c(\Gáivuotna\, \Kaivuono\, \Kåfjord\)",
               "c(\Hamarøy\, \Hábmer\)",
               "c(\Harstad\, \Hárstták\)",
               "Hemne",
               "Herøy i møre og romsdal",
               "Herøy i nordland",
               "Eide",
               "Flora",
               "c(\Lavangen\, \Loabák\)",
               "Vikna",
               "c(\Nordreisa\, \Raisi\, \Ráisa\)",
               "Orkdal",
               "Lenvik",
               "Sør-Varanger",
               "c(\Sortland\, \Suortá\)",
               "Eid",
               "c(\Nesseby\, \Unjárga\)"
)
)
```


