---
title: "Maximal allowed biomass per aquaculture location"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
date: "`r Sys.Date()`"
output:
  html_document:
    css: '~/github/coastrisk/src/style_coast.css'
    toc: true
    toc_float: true
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: 
  pdf_document:
    toc: true
---

```{r setup, include=FALSE, }
knitr::opts_chunk$set(echo = TRUE, message=FALSE)
source('~/github/coastrisk/src/common.r')
```

```{r additional libraries, inlcude = FALSE}
library(data.table)
library(vroom)
library(rebus)
library(here)
```

Here i will prepare a table on MTB (maximal tillat biomasse) per aquaculture production site.
Then i will join this table with production areas, for use in CoastRisk project.


# Load aquaculture registry data (provided by FISKEDIR in May 2019). There is also no data for year 2016
```{r}
year2010 <-read_excel(file.path("/Volumes/ftp.imr.no/Aquaculture/Akvakultur_tillatelser_2010_2018/2010-06-28 Register-internett.xlsx"), skip = 1)
year2011 <-read_excel(file.path("/Volumes/ftp.imr.no/Aquaculture/Akvakultur_tillatelser_2010_2018/2011-01-03 Register-internett.xlsx"), skip = 1)
year2012 <-read_excel(file.path("/Volumes/ftp.imr.no/Aquaculture/Akvakultur_tillatelser_2010_2018/2012-12-31 Register-internett.xlsx"),skip = 1)
year2013 <-read_excel(file.path( "/Volumes/ftp.imr.no/Aquaculture/Akvakultur_tillatelser_2010_2018/2013-12-30 Akvakulturregisteret.xlsx"),skip = 1)
year2014 <-read_excel(file.path( "/Volumes/ftp.imr.no/Aquaculture/Akvakultur_tillatelser_2010_2018/2015-01-05 Akvakulturregisteret.xlsx"),skip = 1)
year2015 <-read_excel(file.path("/Volumes/ftp.imr.no/Aquaculture/Akvakultur_tillatelser_2010_2018/2016-01-11 Akvakulturregisteret.xlsx"),skip = 1)
year2017 <-read_excel(file.path("/Volumes/ftp.imr.no/Aquaculture/Akvakultur_tillatelser_2010_2018/2018-01-01 Akvakulturregisteret.xlsx"),skip = 1)
year2018 <-read_excel(file.path("/Volumes/ftp.imr.no/Aquaculture/Akvakultur_tillatelser_2010_2018/2018-12-31 Akvakulturregisteret.xlsx"),skip = 1)

```

I need to homogenize data a bit before merging

```{r}
year2010<-year2010 %>% mutate(TILL_KAP = as.numeric(TILL_KAP),
                              LOK_KAP = as.numeric(LOK_KAP))


tillatelser2010_18 <- bind_rows(year2010, year2011, year2012, year2013, year2014, year2015, year2017, year2018) %>%
  mutate(Year = c(
    rep(2010, nrow(year2010)),
    rep(2011, nrow(year2011)),
    rep(2012, nrow(year2012)),
    rep(2013, nrow(year2013)),
    rep(2014, nrow(year2014)),
    rep(2015, nrow(year2015)),
    rep(2017, nrow(year2017)),
    rep(2018, nrow(year2018))
  )) %>%
  select(27, c(2:26))
         
```

# Filter out North-Norwegian municipalities
```{r}
tillatelser2010_18_nord <- tillatelser2010_18 %>% filter(TILL_KOMNR >= 1800 & TILL_KOMNR <= 2100)
```


# Clean the licences data
What are the production types that are in the table?
We will need only the rows related to salmon and trout since only these two types of aquacutlure are included in the project
```{r}
unique(tillatelser2010_18$PRODUKSJONSFORM)
```

But i can also sort by species, it is simpler
```{r}
unique(tillatelser2010_18$ART)
```

```{r}
str_view(tillatelser2010_18$ART, pattern = "ørret", match = T)#Regnbueørret
str_view(tillatelser2010_18$ART, pattern = "Ørret", match = T)#Ørret
```

```{r}
str_view(tillatelser2010_18$ART, pattern = "Laks", match = T)#Laks
```

Filter the rows for salmon and trout.
NB! The table includes main production farms  but also stamfisk and settesfisk

```{r}
mtb_prep <- tillatelser2010_18 |> 
  filter(ART %in% c("Laks", "Regnbueørret", "Ørret"))
```


Further data cleaning: remove columns that we don't need, rename columns to English

```{r}
mtb_prep2 <- mtb_prep |> 
  select(-c(
            NAVN,
            ADRESSE,
            POSTNR,
            POSTSTED,
            TILDELINGSTIDSPUNKT,
            TIDSBEGRENSET,
            TILL_KOMNR,
            TILL_KAP,
            TILL_KOM,
            FORMÅL,
            TILL_ENHET,
            LOK_PLASS,
            UTGÅR_DATO)) |> 
  rename(year = Year,
         production_form = PRODUKSJONSFORM,
         species = ART,
         location_number = LOK_NR,
         location_name = LOK_NAVN,
         municip_number = LOK_KOMNR,
         municip = LOK_KOM,
         environment = VANNMILJØ,
         location_capacity = LOK_KAP,
         units_of_capacity = LOK_ENHET,
         W_GEOWGS84 =  Ø_GEOWGS84
         ) |> 
  select(-2)
```

# Prepare production areas data

```{r}
mcp_prod <- read.csv(file.path("./data/raw/municipality_productionarea.csv"),sep = " ")
```

```{r}
mcp_prod_prep <- mcp_prod %>% 
  rename(county_number = fylkenummer,
         municip_number = kommunenummer,
         municip_name = komname,
         total_mcp_area = Totalt_areal_inkl_12nm,
         sea_mcp_area = Havflate_12nm,
         prod_area_id = id,
         prod_area_name = name
         ) %>% 
  select(-knum) %>% 
  mutate(
    prod_area_name = str_replace_all(
      prod_area_name, pattern = "->", replacement = "-"
    )
  ) %>% 
  mutate(
    prod_area_name = str_replace_all(
      prod_area_name, pattern = fixed("+"), replacement = "-"
    )
  )

mcp_prod_prep2 <- mcp_prod_prep %>% 
  mutate(municip_number =
           ifelse(
             str_count(municip_number, pattern = regex("[:digit:]")) == 3,
                       str_c("0",municip_number),
                       municip_number
             )
         )
```


# Merge mtb table and production areas

```{r}
mtb_prep3 <- mtb_prep2 |> 
left_join(mcp_prod_prep2[,c(
    "municip_number",
    "municip_name",
    "prod_area_id",
    "prod_area_name"
  )],
  by = "municip_number")
```

There were many occasions when production areas did not match municipalities in mtb table
```{r}
nonmatches <- mtb_prep3 |> 
      filter(is.na(prod_area_name)) |> 
                       group_by(municip_number, municip) |> 
                       tally()

nonmatches
      
```

Bjarkoy municipality merged with Harstad in 2013, so i will just add a new number 1903 and name Harstad. Later, if we want to aggregate all capacities per municipalities, we will just take a sum of capacities in Harstand and Bjarkoy


I will make changes in the municipalities' number in the mtb table and will merge again with production areas
I replace also municipalities' numbers for Trondelad, they are old in the mtb table but new in production areas table

Changes in municipalities' numbers table
```{r}
convert_mcp <- read_excel(file.path("../administrative/kommune_endringer_2000_2018.xlsx"))

convert_mcp_prep1 <-convert_mcp %>% 
  mutate(
    municip_number_old = str_trim(municip_number_old)) %>% 
   mutate(
     municip_number_new = str_trim(municip_number_new)) 
```





Change to a new number: Roros and Rissa (merged with Indre Fossen)
```{r}
mtb_prep4 <- mtb_prep2 |> 
  mutate(municip_number = replace(municip_number, municip_number == "1915", "1903")) |> 
  mutate(municip_number = replace(municip_number, municip_number == "1901", "1903")) |> 
  mutate(municip_number = replace(municip_number, municip_number == "1640", "5025")) |> 
  mutate(municip_number = replace(municip_number, municip_number == "1624", "5024"))
```


```{r}
foo_trond <- mtb_prep4 |> 
  filter(municip_number >=1600 & municip_number < 1800) |> 
  left_join(convert_mcp_prep1[, c(1,3)], 
            by = c("municip_number" = "municip_number_old")) |> 
  mutate(municip_number = municip_number_new) |> 
  select(-municip_number_new)
```

Merge Trondelag municipalities with others

```{r}
mtb_prep5 <- mtb_prep4 |> 
  filter(municip_number >= 1800 | municip_number < 1600) 

mtb_prep6 <- bind_rows(mtb_prep5, foo_trond) |> 
  left_join(mcp_prod_prep2[,c(
    "municip_number",
    "municip_name",
    "prod_area_id",
    "prod_area_name"
  )],
  by = "municip_number")
```

And check non-matches again
```{r}
sum(is.na(mtb_prep6$prod_area_name))

nonmatches2 <- filter(mtb_prep6, is.na(prod_area_name)) |> 
  group_by(municip, municip_number) |> 
  tally()

nonmatches2
```

These are inland municipalities that do not have matches in production areas table, not sure why. I will leave these rows without changes


```{r}
mtb_final <- mtb_prep6 |> 
  select(-municip) |> 
  rename(municip = municip_name) |> 
  select(year, municip_number, municip, everything())
```







# Save the final table
```{r}
write.csv(mtb_final, "./data/output/locations_capacity_2010_2018.csv")
```


# Table on MATFISK with volume-measured capacity

The table above can be used to see how many production sites where there each year, but capacity cannot be compared between the areas - sometimes the measurement is tonn, sometines count and m3. I can only keep m3, kg and tonn-units observations, if we want to present the capacity as a comparable numeric variable.

Transform kg to tonn and m3 to tonn. I will only keep MATFISK because, usually, SETTEFISK and STAMFISK location capacities are measured as counts. Also keep only the observations where capacity is measure as tonn, kg or m3.

Conversion from m3 to tonns: ant m3 x 0,065 = ant ton

Note that some capacities ara zero, for instance, location 	13189 in 2014.
```{r}
mtb_prep7 <- mtb_final |> 
  filter(production_form == "MATFISK") |> 
  filter(units_of_capacity %in% c("KG", "TN", "M3")) |> 
  mutate(location_capacity = ifelse(units_of_capacity == "KG",
         location_capacity/1000,
         location_capacity)) |> 
  mutate(location_capacity = ifelse(units_of_capacity == "M3",
         location_capacity*0.065,
         location_capacity))
```

# Save the updated table

```{r}
write.csv(mtb_prep7, "./data/output/locations_capacity_marfisk_only_2010_2018.csv")
```

