---
title: "Plotting_aquaculture_production"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output:
  html_document:
    css: '~/github/coastrisk/src/style_coast.css'
    toc: true
    toc_float: true
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: 
  pdf_document:
    toc: true
---

```{r setup, include=FALSE, }
knitr::opts_chunk$set(echo = TRUE, message=FALSE)
source('~/github/coastrisk/src/common.r')
```

```{r}
library(sf)
library(mapdata)
library(maptools)
library(rgdal)
library(rgeos)
library(maps)
library(rnaturalearth)
library(rnaturalearthdata)
library(stringr)
```

# Loading data: aquaculture yield per region and regions geometry
```{r}
aqua_prod <-read.csv(file.path("./data/annual_aqua_production_sum_per_prodregion.csv"))
```

Check for which prod_area*year combination we do not have aquaculture yield value: Ost-Finnmark in 2012. There, the total annual aquaculture yield was negative, and I considered that as missing data (we don't know the real yield value).
```{r}
filter(aqua_prod, is.na(production_per_region))
```


```{r}
prod_areas <- readOGR(dsn = "/Volumes/ftp.imr.no/ProductionAreas/OGRgeoJSON_prodomr")
```

Check the coordinate reference system of the production areas shape file
```{r}
crs(prod_areas)
```

# Plotting aquaculture production per year and produciton region
## Create a theme for aquaculture production plots
```{r}
theme_akva <- function(...) {
  theme_hc() +
    theme(
      axis.text.x = element_text(
        size = 12, angle = 90, hjust = 1, vjust = 0.5,
        face = "italic", color = "grey51"
      ),
      axis.text.y = element_text(size = 10, angle = 0, face = "italic", color = "grey51"),
      axis.title.y = element_text(margin = margin(t = 0, r = 15, b = 0, l = 0), face = "italic", size = 12, color = "grey22"),
      plot.title = element_text(hjust = 0.5, vjust = 0.5),
      axis.ticks.x = element_line(size = 0.2),
      axis.ticks.length = unit(0.1, "cm"),
      #axis.line.x = element_line(size = 0.3, color = "black"),
      strip.text = element_text(size = 12, face = "italic", color = "grey22"),
      strip.background = element_rect(fill = "white", color ="white"),
      panel.background = element_rect(fill = "white"),
      plot.background = element_rect(fill = "white"),
      panel.grid.major.y = element_line(color = "grey", size = 0.2),
      panel.grid.major.x = element_line(color = "grey", size = 0.2)
     
    )
}
```


Prepare x-axis with years and the breaks between them
```{r}
breaks1 <- seq(2000, 2020, by = 1)
labels1 <- as.character(breaks1)
labels1[!breaks1 %% 2 == 0] <- ""

breaks2 <-seq(0,250000,by = 50000)
labels2 <- imap_chr(breaks2, function(., id){
  return(paste0(formatC(breaks2[id], format="f", big.mark="," , digits = 0),
                             ""))
})
```

## Prepare the aquaculture data for plotting
The production values are in kg, I will divide them all by 1000 and present as tons.
```{r}
aqua_prod_prep2 <- aqua_prod %>% 
  mutate(production_tons = production_per_region/1000)
```



```{r}
plot1 <-ggplot(
  data = aqua_prod_prep2,
  mapping = aes(x = year,
                y = production_tons)
) +
  geom_line(color = "mediumblue", size = 1.1, alpha = 0.7) +
  facet_wrap(prod_area_name ~ .) +
  scale_x_continuous(limits = c(2000, 2020), breaks = breaks1, labels = labels1) +
  scale_y_continuous(breaks = breaks2, labels = labels2) +
  labs(y = "Annual aquaculture yield, tons", x = "") +
  geom_hline(yintercept = 0, color = "black", size = 0.3) +
  theme_akva()
```


```{r fig.width = 4, fig.height = 3, dpi = 300}
plot1
```
```{r}
ggsave("figs/aqua_yield_per_prodregion.pdf", plot = plot1, width = 10, height =  8, dpi = 300)
```

# Mapping aquacultlure yield per produciton reigon


```{r}
prodareas_sf <- st_as_sf(prod_areas, coords= c("long", "lat"), agr="identity")
```

Now we need to merge the table of geometries of production regions and the aquaculture yield tables. The names of production regions are written differently in the two tables, so I have to homogenize them first.

```{r}

prodareas_sf_prep2 <- prodareas_sf %>% 
   mutate(
    prod_area_name = str_replace_all(
      name, pattern = "->", replacement = "-"
    )
  ) %>% 
  mutate(
    prod_area_name = str_replace_all(
      prod_area_name, pattern = fixed("+"), replacement = "-"
    )
  ) %>% 
  select(-name)
```

I merge the two tables and transform production from kg to tons:
```{r}
plot_aquaprod <- left_join(
  aqua_prod,
  prodareas_sf_prep2, 
  by = "prod_area_name") %>% 
  mutate(production_tons = production_per_region/1000) %>% 
  select(-id) %>% 
  st_as_sf(.)
```




## Map aquaculture production 

```{r}
#https://gotellilab.github.io/GotelliLabMeetingHacks/NickGotelli/ColorPalettes.html

addalpha <- function(colors, alpha=0.7) {
  r <- col2rgb(colors, alpha=T)
  # Apply alpha
  r[4,] <- alpha*255
  r <- r/255.0
  return(rgb(r[1,], r[2,], r[3,], r[4,]))
}

cols <- addalpha(brewer.pal(5, "Purples"))
```


```{r}
world <- ne_countries(scale = "medium", returnclass = "sf")

```


I will use the same breaks of production values as for line chart, but I will make them as intervals.
```{r}
#50000, 100000, 150000, 200000
breaks <-seq(0,250000,by = 50000)
labels <- imap_chr(breaks, function(., id){
  return(paste0(formatC(breaks[id], format="f", big.mark="," , digits = 0),
                             "",
                            " – ",
                             formatC(breaks[id + 1], format="f", big.mark="," , digits = 0),
                             ""))
})

labels <- labels[1:length(labels) - 1] #5 intervals in total
```

There is one value of production in tons which fall out of the bound - 250452, i will replace it with 250000, because for the aesthetics of the map it is better (and more precise values can be read from line chart).
```{r}
plot_aquaprod_prep2  <- plot_aquaprod %>% 
  filter(!is.na(production_tons)) %>% 
  mutate(production_tons = 
           replace(production_tons,
                   production_tons >= 250000,
                   250000)) %>% 
  mutate(prod_quantiles = cut(production_tons,
                               breaks = breaks,
                               labels = labels,
                               include.lowest = T))
```


```{r}
base_map <- ggplot() +
  geom_sf(
   data = world, 
   fill = "antiquewhite1", 
   color = "dimgray", 
   size = 0.3) +
  theme(axis.text.x = element_text(size = 10,  color = "grey51"),
        axis.text.y = element_text(size = 10,  color = "grey51"),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 15,  face = "bold"),
        panel.background = element_rect(fill = "aliceblue") ,
        panel.grid.major = element_line(color = "#dbdbd9", linetype = "dashed", size = 0.2),
        plot.title = element_text(hjust = 0.5, vjust = 0.5, size = 17, face = "bold"),
        strip.text = element_text(size = 13, face = "italic", color = "grey24"),
        strip.background = element_rect(fill = "white", color ="white"),
  ) +
  scale_x_continuous(breaks = c(5,15,25)) +
  geom_sf(
    data = plot_aquaprod_prep2,
    mapping = aes(fill = prod_quantiles),
    color = "white",
    size = 0.2
  ) +
  coord_sf(xlim = c(2, 34), ylim = c(55, 72), expand = FALSE) +
  labs(x = "Longitude", y = "Latitude") +
  scale_fill_manual(
    values = cols,
    name = "Aquaculture yield, tons",
    guide = guide_legend(
      keyheight = unit(9, units = "mm"),
      title.position = "top"
    )
  ) +
  facet_wrap(year ~.) +
  labs(
    x = NULL,
    y = NULL
  ) 

```

Add the id of each production region
```{r}
prodareas_names <- cbind(plot_aquaprod_prep2,st_coordinates(st_centroid(plot_aquaprod_prep2)))
```

```{r}
main_map <-base_map +
  geom_sf_text(data = prodareas_names, aes(X,Y, label = prod_area_id),fontface = "bold", size = 3) 
```

```{r, fig.width=10, fig.height=7}
main_map
```
```{r}
ggsave("./figs/aqua_yield_maps_per_prodarea.tiff", main_map, width = 12, height = 14, dpi = 300)

```



