---
title: "Plotting_aquaculture_production"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output:
  html_document:
    css: '~/github/coastrisk/src/style_coast.css'
    toc: true
    toc_float: true
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: 
  pdf_document:
    toc: true
---

```{r setup, include=FALSE, }
knitr::opts_chunk$set(echo = TRUE, message=FALSE)
source('~/github/coastrisk/src/common.r')
```

```{r}
library(sf)
library(mapdata)
library(maptools)
library(rgdal)
library(rgeos)
library(maps)
library(rnaturalearth)
library(rnaturalearthdata)
```

# Loading data: aquaculture yield per region and regions geometry
```{r}
aqua_prod <-read.csv(file.path("./data/annual_aqua_production_sum_per_prodregion.csv"))
```

```{r}
prod_areas <- readOGR(dsn = "/Volumes/ftp.imr.no/ProductionAreas/OGRgeoJSON_prodomr")
```
Check the coordinate reference system of the production areas shape file
```{r}
crs(prod_areas)
```

# Plotting aquaculture production per year and produciton region
## Create a theme for aquaculture production plots
```{r}
theme_akva <- function(...) {
  theme_hc() +
    theme(
      axis.text.x = element_text(
        size = 12, angle = 90, hjust = 1, vjust = 0.5,
        face = "italic", color = "grey51"
      ),
      axis.text.y = element_text(size = 10, angle = 0, face = "italic", color = "grey51"),
      axis.title.y = element_text(margin = margin(t = 0, r = 15, b = 0, l = 0), face = "italic", size = 12, color = "grey22"),
      plot.title = element_text(hjust = 0.5, vjust = 0.5),
      axis.ticks.x = element_line(size = 0.2),
      axis.ticks.length = unit(0.1, "cm"),
      #axis.line.x = element_line(size = 0.3, color = "black"),
      strip.text = element_text(size = 12, face = "italic", color = "grey22"),
      strip.background = element_rect(fill = "white", color ="white"),
      panel.background = element_rect(fill = "white"),
      plot.background = element_rect(fill = "white"),
      panel.grid.major.y = element_line(color = "grey", size = 0.2),
      panel.grid.major.x = element_line(color = "grey", size = 0.2)
     
    )
}
```


Prepare x-axis with years and the breaks between them
```{r}
breaks1 <- seq(2000, 2020, by = 1)
labels1 <- as.character(breaks1)
labels1[!breaks1 %% 2 == 0] <- ""

breaks2 <-seq(0,250000,by = 50000)
labels2 <- imap_chr(breaks2, function(., id){
  return(paste0(formatC(breaks2[id], format="f", big.mark="," , digits = 0),
                             ""))
})
```

## Prepare the aquaculture data for plotting
The production values are in kg, I will divide them all by 1000 and present as tons.
```{r}
aqua_prod_prep2 <- aqua_prod %>% 
  mutate(production_tons = production_per_region/1000)
```



```{r}
plot1 <-ggplot(
  data = aqua_prod_prep2,
  mapping = aes(x = year,
                y = production_tons)
) +
  geom_line(color = "mediumblue", size = 1.1, alpha = 0.7) +
  facet_wrap(prod_area_name ~ .) +
  scale_x_continuous(limits = c(2000, 2020), breaks = breaks1, labels = labels1) +
  scale_y_continuous(breaks = breaks2, labels = labels2) +
  labs(y = "Annual aquaculture yield, tons", x = "") +
  geom_hline(yintercept = 0, color = "black", size = 0.3) +
  theme_akva()
```


```{r fig.width = 4, fig.height = 3, dpi = 300}
plot1
```
```{r}
ggsave("figs/aqua_yield_per_prodregion.pdf", plot = plot1, width = 10, height =  8, dpi = 300)
```




3) Plot a production in year 2005 and 2018 to compare, per region.