---
title: "Plotting aquaculture production regions"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output:
  html_document:
    css: '~/github/coastrisk/src/style_coast.css'
    toc: true
    toc_float: true
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: 
  pdf_document:
    toc: true
---


```{r setup, include=FALSE, }
knitr::opts_chunk$set(echo = TRUE, message=FALSE)
source('~/github/coastrisk/src/common.r')
```

```{r inlcude=FALSE}
library(sf)
library(mapdata)
library(maptools)
library(rgdal)
library(rgeos)
library(maps)
library(rnaturalearth)
library(rnaturalearthdata)
```

# Preparing data for mapping
Here i want to plot the aquaculture production regions, to use this map in the coastrisk paper 1. There will be no more other information on the map, just the regions with their id numbers.

```{r}
prod_areas <- readOGR(dsn = "/Volumes/ftp.imr.no/ProductionAreas/OGRgeoJSON_prodomr")
```

```{r}
prodareas_sf <- st_as_sf(prod_areas, coords= c("long", "lat"), agr="identity")
```

I will first make the names of the regions cleaner, and in the same way they are written in aquaculture maps

```{r}
prodareas_sf_prep2 <- prodareas_sf %>% 
   mutate(
    prod_area_name = str_replace_all(
      name, pattern = "->", replacement = "-"
    )
  ) %>% 
  mutate(
    prod_area_name = str_replace_all(
      prod_area_name, pattern = fixed("+"), replacement = "-"
    )
  ) %>% 
  select(-name)
```


# Mapping the production regions
```{r}
world <- ne_countries(scale = "medium", returnclass = "sf")
```

```{r}
base_map <- ggplot() +
  geom_sf(
   data = world, 
   fill = "antiquewhite1", 
   color = "dimgray", 
   size = 0.3) +
  theme(axis.text.x = element_text(size = 10,  color = "grey51"),
        axis.text.y = element_text(size = 10,  color = "grey51"),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 15,  face = "bold"),
        panel.background = element_rect(fill = "aliceblue") ,
        panel.grid.major = element_line(color = "#dbdbd9", linetype = "dashed", size = 0.2),
        plot.title = element_text(hjust = 0.5, vjust = 0.5, size = 17, face = "bold")
  ) +
  scale_x_continuous(breaks = c(5,10,15,20, 25,30)) +
  geom_sf(
    fill = NA,
    data = prodareas_sf_prep2,
    color = "midnightblue",
    size = 0.8
  ) +
  coord_sf(xlim = c(2, 34), ylim = c(55, 75), expand = FALSE) +
  labs(x = "Longitude", y = "Latitude") +
  labs(
    x = NULL,
    y = NULL
  ) 

```

Adding id's of production region superimposed on the map. The 1st region does not plot the label very well, it is plotted on the border with region 2.
I will manually adjust the coordinates of a centroid for id.
```{r}
prodareas_names <- cbind(prodareas_sf_prep2,st_coordinates(st_centroid(prodareas_sf_prep2)))

prodareas_names_prep2 <- prodareas_names %>% 
  st_set_geometry(., NULL) %>% 
  mutate(Y = replace(
    Y,
    prod_area_name == "Svenskegrensen - Jæren",
    58.2
  )) %>% 
  st_as_sf(.,  coords = c("X", "Y"), crs = 4326, remove = FALSE, agr = "identity")

```


```{r}
main_map <- base_map +
  geom_sf_text(data = prodareas_names_prep2,
    aes(X,Y,label = id), fontface = "bold", size = 5)
              
```

```{r}
ggsave("./figs/aquaprod_regions.pdf", main_map, width = 6, height = 10, dpi = 300)
```

# Plotting aquaculture production regions and fisheries statistics areas together
loading shapefiles of fisheries statistic areas
```{r}
fish_areas <- readOGR(dsn = "/Volumes/ftp.imr.no/fisheries_data/fisheries_statistic_areas")
```
Check the coordinate reference system of fisheries areas:they are in UTM, not in long-lat. I will need to transform the coordinates so that production areas and fisheries areas are in the same projection (WGS84, with a code 4326).
```{r}
crs(fish_areas)
```

```{r}
fishareas_sf <- st_as_sf(fish_areas)
fishareas_sf_prep2 <- st_transform(fishareas_sf, crs = 4326)
```

## Map
Base map will be similar to the previous one, but i  will extend the coordinates.

```{r}
base_map2 <- ggplot() +
  geom_sf(
   data = world, 
   fill = "antiquewhite1", 
   color = "dimgray", 
   size = 0.3) +
  theme(axis.text.x = element_text(size = 10,  color = "grey51"),
        axis.text.y = element_text(size = 10,  color = "grey51"),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 15,  face = "bold"),
        panel.background = element_rect(fill = "aliceblue") ,
        panel.grid.major = element_line(color = "#dbdbd9", linetype = "dashed", size = 0.2),
        plot.title = element_text(hjust = 0.5, vjust = 0.5, size = 17, face = "bold")
  ) +
  scale_x_continuous(breaks = c(0,5,10,15,20, 25,30)) +
  geom_sf(
    fill = NA,
    data = prodareas_sf_prep2,
    color = "midnightblue",
    size = 0.7
  ) +
  coord_sf(xlim = c(0, 34), ylim = c(55, 75), expand = FALSE) +
  labs(x = "Longitude", y = "Latitude") +
  labs(
    x = NULL,
    y = NULL
  ) 

```

Adding fisheries statistics areas on the top of aquaculture production areas
```{r}
main_map2 <- base_map2 + geom_sf(
    fill = NA,
    data = fishareas_sf_prep2,
    color = "darkred",
    size = 0.3,
    linetype = "dotted"
  ) +
   coord_sf(xlim = c(0, 34), ylim = c(55, 75), expand = FALSE)
  
```

Save this version of the map - without any labels
```{r}
ggsave("./figs/fishareas_n_aquaprod_regions.pdf", main_map2, width = 8, height = 9, dpi = 300)
```


Adding the names of regions and fisheries polygons (although, they won't be very clear - too many polygons). In the same way as for production areas - calculate centroids and bind them to sf dataframe.
```{r}
fishareas_names <- cbind(fishareas_sf_prep2,st_coordinates(st_centroid(fishareas_sf_prep2)))
```
```{r}
main_map3 <-
  main_map2 +
  geom_sf_text(data = prodareas_names_prep2,
    aes(X,Y,label = id), fontface = "bold", size = 4, color = "midnightblue") +
  geom_sf_text(data = fishareas_names,
               aes(X,Y, label = Lokasjon),fontface = "bold", size = 2, color = "darkred" )
  
```

```{r}
ggsave("./figs/fishareas_n_aquaprod_regions_labelled.pdf", main_map3, width = 8, height = 9, dpi = 300)
```