---
title: "Defining intersecitons between counties and produciton areas"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output:
  html_document:
    css: '~/github/coastrisk/src/style_coast.css'
    toc: true
    toc_float: true
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: 
  pdf_document:
    toc: true
---

```{r setup, include=FALSE, }
knitr::opts_chunk$set(echo = TRUE, message=FALSE)
source('~/github/coastrisk/src/common.r')
```

```{r}
library(sf)
library(mapdata)
library(maptools)
library(rgdal)
library(rgeos)
library(maps)
library(rnaturalearth)
library(rnaturalearthdata)
library(stringr)
```


Here I need to define the intersection between counties polygons and production areas polygons. Often, counties will be larger than prod areas, so several prod areas can fit in one county. 
Later  I need to know the area of intersection between each prod area and a county, and calculate the proportion of each prod area within a county. Using this proporiton, I can calculate the number of waste water stations per production area.

```{r}
prod_areas <- readOGR(dsn = "/Volumes/ftp.imr.no/ProductionAreas/OGRgeoJSON_prodomr")
prodareas_sf <- st_as_sf(prod_areas, coords= c("long", "lat"), agr="identity")
```

Polygons of Norwegian fylker

https://gis.stackexchange.com/questions/151613/reading-feature-class-in-file-geodatabase-using-r

```{r}
require(rgdal)
fylker <-readOGR(dsn = "/Users/marinaespinasse/github/nor-prep/prep/administrative/raw/shapefiles/NO_Fylker")
fylkersf <- st_as_sf(fylker,coords= c("long", "lat"), crs = 4326)


```



```{r}
path = "/Users/marinaespinasse/Downloads/Basisdata_0000_Norge_25833_Fylker_FGDB.gdb"
fylker <- st_read(path)
```
Check the projection of both files - seems to be WGS84 for both

```{r}
 st_crs(fylker)
 st_crs(prod_areas)
 sf_use_s2(FALSE)
```

Extract counties' names from fylker table

https://www.eriksmistad.no/norges-fylker-og-kommuner-i-geojson-format/

```{r}
fylker$name <- c("Rogaland", "Vestfold og Telemark","Nordland",
                 "Agder", "Innlandet","Møre og Romsdal", "Vestland",
                 "Oslo", "Viken", "Trøndelag", "Troms og Finnmark")

fylker_prep <- fylkersf |> 
 rename(county_nubmer = NR,
         county_name = NAVN) |> 
 select(-c(OBJTYPE, OBJECTID)) |> 
  clean_names()
```

# Calcualte intersections between produciton areas and counties


Make fylker crs same as crs of production areas

```{r}
fylker_trans <- st_transform(fylker_prep, crs = st_crs(prodareas_sf))
#fylker_trans2 = st_transform(fylker_prep, crs = "+proj=longlat +datum=WGS84")
```

```{r}
intersects <- st_intersection(prodareas_sf, fylker_trans)
intersects$intersection_area_m2 <-as.vector(st_area(intersects))


fylker_trans_centroids <- cbind(fylker_trans, st_coordinates(st_centroid(fylker_trans)))
```


```{r}
fylker_n_pa <- ggplot() +
geom_sf(data = fylker_trans, fill = NA, color = "darkblue") +
geom_sf(data = prodareas_sf, fill=NA, color = "darkred", size = 1)  +
  geom_sf_text(data = fylker_trans_centroids, aes(X,Y, label = county_nubmer), fontface = "bold", size = 3, color = "green") +
  theme_bw() 
```


```{r}
ggsave("figs/counties_n_prodareas_map.pdf", plot = fylker_n_pa, width = 7, height =  10, dpi = 300 )
```

As the plot suggest, some counties do not have any intersection with production areas.

Clean the table on intersections between counties and produciton areas

```{r}
intersects_final <- intersects |> 
  rename(prod_area_id = id,
         prod_area_name = name) |> 
  select(prod_area_id, prod_area_name, county_name, county_nubmer, intersection_area_m2) |> 
  st_drop_geometry()
```

```{r}
write_csv(intersects_final, file = "./data/counties_prodareas_intersections.csv")
```

