---
title: "Standardized population size per PA "
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
date: "`r Sys.Date()`"
output:
  html_document:
    css: '~/github/coastrisk/src/style_coast.css'
    toc: true
    toc_float: true
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: 
  pdf_document:
    toc: true
---

```{r setup, include=FALSE, }
knitr::opts_chunk$set(echo = TRUE, message=FALSE)
source('~/github/coastrisk/src/common.r')
```

```{r additional libraries, inlcude = FALSE}
library(data.table)
library(vroom)
library(rebus)
library(here)
library(sf)
library("ggsci")
```


Here I will calculate the population size for each production area, based on 2019 data. Then I will standardize the estiamtes by maximal observed population (taking max as 1, and the rest - as proportions to it).

This is the table of population per proeduciton area, where Anna used 5 by 5 km grid size of population densities, and she also applied a 50 km buffer from the coastline (to focus on the coastal population only).

```{r}
population_50km <- read_table("/Volumes/ftp.imr.no/Tourism&Recreation/Population5kmGrid2001-19/square5_buffer50_all_pop0119PA.txt")

prod_areas <- read_csv(here("prep", "aquaculture", "data", "output", "prod_areas.csv"))
```
```{r}
pop_st <- population_50km |> 
 clean_names() |> 
  filter(year == "2019") |> 
  left_join(prod_areas, by = c("id" = "prod_area_id")) |> 
  mutate(standardized_population = pop_tot/max(pop_tot)) |> 
  rename(prod_area_id = id)
```

```{r}
write_csv(pop_st, file = "./data/output/standardized_population_50km_by_maximum.csv")
```

