---
title: "Edge bunding diagram for paper 5"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
date: "`r Sys.Date()`"
output:
  html_document:
    css: '~/github/coastrisk/src/style_coast.css'
    toc: true
    toc_float: true
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: 
  pdf_document:
    toc: true
---

```{r setup, include=FALSE, }
knitr::opts_chunk$set(echo = TRUE, message=FALSE)
source('~/github/coastrisk/src/common.r')
```

```{r additional libraries, inlcude = FALSE, message=FALSE}
library(data.table)
library(vroom)
library(rebus)
library(here)
library(ggraph)
library(igraph)
library(tidyverse)
library(RColorBrewer)
```

# Load the data

These are the edges table - sectors, pressures and ecosystems and the next is connections - how they are connected to each other

```{r}
edges <- read.csv("./data/edges.csv")
connect <- read.csv("./data/connections.csv")
```

```{r}
connect_prep <- connect |> 
  select(c(From, To)) |> 
  rename(from = From,
         to = To)

```


# Prepare the main table of connections

I am adding a fake value column, just to try. I can remove it later and add "real" values for circles (scaled or standardized values)
```{r}
myvertices  <-  data.frame(
  name = unique(c(as.character(edges$from), as.character(edges$to))) ,
  value = runif(59)
) 
```


The group column specifies where the vertex comes from, or a closest higher level. For instance, origin will have NA, while sectors and pressures will have origin as a group (as closest next level above).

```{r}
myvertices$group  <-  edges$from[ match( myvertices$name, edges$to ) ]
```

More attributes for plotting.
ID is a number of a component except the components that are defining classification (sector, pressure, ecosystem). So everything else are the components (called subgroups in the example codes)
```{r}
myvertices$id <- NA
myleaves <- which(is.na( match(myvertices$name, edges$from) ))
mynleaves <- length(myleaves)
myvertices$id[ myleaves ] <- seq(1:mynleaves)
```

```{r}
# Create a graph object
mygraph <- igraph::graph_from_data_frame( edges, vertices=myvertices )
 
# The connection object must refer to the ids of the leaves:
from  <-  match( connect_prep$from, myvertices$name)
to  <-  match( connect_prep$to, myvertices$name)
```


```{r, fig.width=9, fig.height=9}
ggraph(mygraph, layout = 'dendrogram', circular = TRUE) + 
  geom_conn_bundle(data = get_con(from = from, to = to), alpha=0.2, width=0.9, aes(colour=..index..)) +
  scale_edge_colour_distiller(palette = "Blues") +
  
  geom_node_text(aes(x = x*1.15, y=y*1.15, filter = leaf, label=name, 
                     angle = pmax((acos(x/(((x^2)+(y^2))^0.5))*sign(y)*180/pi-((sign(x)==-1)*(sign(y)==1)*180)+((sign(x)==-1)*(sign(y)==-1)*180)),-180,na.rm=TRUE), 
                     hjust=(sign(x)==-1), 
                     colour=group), size=3.5, alpha=1) +
  
  geom_node_point(aes(filter = leaf, x = x*1.07, y=y*1.07, colour=group, size=value, alpha=0.2)) +
  scale_colour_manual(values= rep( brewer.pal(5,"Dark2") , 30)) +
  scale_size_continuous( range = c(0.1,10) ) +
  
  theme_void() +
  theme(
    legend.position="none",
    plot.margin=unit(c(0,0,0,0),"cm"),
  ) +
  expand_limits(x = c(-1.3, 1.3), y = c(-1.3, 1.3))
```

```{r}
ggsave("./edge_bunding_p5.pdf", width = 9, height = 9, dpi = 300)
```

