---
title: "Trial of edge bunding diagram"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
date: "`r Sys.Date()`"
output:
  html_document:
    css: '~/github/coastrisk/src/style_coast.css'
    toc: true
    toc_float: true
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: 
  pdf_document:
    toc: true
---

```{r setup, include=FALSE, }
knitr::opts_chunk$set(echo = TRUE, message=FALSE)
source('~/github/coastrisk/src/common.r')
```

```{r additional libraries, inlcude = FALSE, message=FALSE}
library(data.table)
library(vroom)
library(rebus)
library(here)
library(ggraph)
library(igraph)
library(tidyverse)
library(RColorBrewer)
```

Here  i will just try to plot edge bunding diagram using the data from online examples. Then, i will try to create my own, but fictional data, sth similar to coastrisk paper 5, and will plot it as well.

# Without g the labels
## create a data frame giving the hierarchical structure of your individuals
```{r}
set.seed(1234)
d1 <- data.frame(from="origin", to=paste("group", seq(1,10), sep=""))
d2 <- data.frame(from=rep(d1$to, each=10), to=paste("subgroup", seq(1,100), sep="_"))
edges <- rbind(d1, d2)
```


 
## create a dataframe with connection between leaves (individuals)
```{r}
all_leaves <- paste("subgroup", seq(1,100), sep="_")
connect <- rbind( 
  data.frame( from=sample(all_leaves, 100, replace=T) , to=sample(all_leaves, 100, replace=T)), 
  data.frame( from=sample(head(all_leaves), 30, replace=T) , to=sample( tail(all_leaves), 30, replace=T)), 
  data.frame( from=sample(all_leaves[25:30], 30, replace=T) , to=sample( all_leaves[55:60], 30, replace=T)), 
  data.frame( from=sample(all_leaves[75:80], 30, replace=T) , to=sample( all_leaves[55:60], 30, replace=T)) )
connect$value <- runif(nrow(connect))
```


 
## create a vertices data.frame. One line per object of our hierarchy
```{r}
vertices  <-  data.frame(
  name = unique(c(as.character(edges$from), as.character(edges$to))) , 
  value = runif(111)
) 
```


## Let's add a column with the group of each name. It will be useful later to color points
The group column specifies where the vertex come from, or a closest  higher level. For instance, origin will have NA, while group1 and group2 will have origin as a group (as closesnt next level above).

```{r}
vertices$group  <-  edges$from[ match( vertices$name, edges$to ) ]
```


# Creating the labels
Next step: computing the label features that will be displayed all around the circle, next to the nodes:

angle → vertical on top and botton, horizontal on the side, and so on.
flip it → labels on the left hand side must be 180° flipped to be readable
alignment → if labels are flipped, they must be right aligned
Those information are computed and added to the vertices data frame.

#Let's add information concerning the label we are going to add: angle, horizontal adjustement and potential flip
#calculate the ANGLE of the labels
```{r}
vertices$id <- NA
myleaves <- which(is.na( match(vertices$name, edges$from) ))
nleaves <- length(myleaves)
vertices$id[ myleaves ] <- seq(1:nleaves)
vertices$angle <- 90 - 360 * vertices$id / nleaves
```


 
## calculate the alignment of labels: right or left
 If I am on the left part of the plot, my labels have currently an angle < -90
```{r}
vertices$hjust <- ifelse( vertices$angle < -90, 1, 0)
```


## flip angle BY to make them readable
```{r}
vertices$angle <- ifelse(vertices$angle < -90, vertices$angle+180, vertices$angle)
vertices$angle2 <-vertices$angle + 90
```

# Plot the labels

```{r  fig.width=7, fig.height=7}
# Create a graph object
mygraph <- igraph::graph_from_data_frame( edges, vertices=vertices )
 
# The connection object must refer to the ids of the leaves:
from  <-  match( connect$from, vertices$name)
to  <-  match( connect$to, vertices$name)
 
# Basic usual argument
ggraph(mygraph, layout = 'dendrogram', circular = TRUE) + 
  geom_node_point(aes(filter = leaf, x = x*1.05, y=y*1.05)) +
  geom_conn_bundle(data = get_con(from = from, to = to), alpha=0.2, colour="skyblue", width=0.9) +
  geom_node_text(aes(x = x*1.1, y=y*1.1, filter = leaf, label=name, angle = angle, hjust=hjust), size=1.5, alpha=1) +
  theme_void() +
  theme(
    legend.position="none",
    plot.margin=unit(c(0,0,0,0),"cm"),
  ) +
  expand_limits(x = c(-1.2, 1.2), y = c(-1.2, 1.2))

```

?? How to change the angle to 180 for each?


## With color
```{r fig.width=7, fig.height=7}
ggraph(mygraph, layout = 'dendrogram', circular = TRUE) + 
  geom_conn_bundle(data = get_con(from = from, to = to), alpha=0.2, width=0.9, aes(colour=..index..)) +
  scale_edge_colour_distiller(palette = "RdPu") +
  
  geom_node_text(aes(x = x*1.15, y=y*1.15, filter = leaf, label=name, angle = angle, hjust=hjust, colour=group), size=2, alpha=1) +
  
  geom_node_point(aes(filter = leaf, x = x*1.07, y=y*1.07, colour=group, size=value, alpha=0.2)) +
  scale_colour_manual(values= rep( brewer.pal(9,"Paired") , 30)) +
  scale_size_continuous( range = c(0.1,10) ) +
  
  theme_void() +
  theme(
    legend.position="none",
    plot.margin=unit(c(0,0,0,0),"cm"),
  ) +
  expand_limits(x = c(-1.3, 1.3), y = c(-1.3, 1.3))
```





# My own example

## create a data frame with structure

Later, you can rename subgroups into pressures and ecosystem parts. But i think, the easiest is to create such a table by hand in Excel, carefully check the connections, and then import it here.

```{r}
set.seed(1234)
myd1 <- data.frame(from="origin", to=paste("sector", seq(1,8), sep=""))
myd2 <- data.frame(from=rep(myd1$to, each=5), to=paste("subgroup", seq(1,40), sep="_"))
myedges <- rbind(myd1, myd2)
```


 
## create a dataframe with connection between leaves (individuals)
```{r}
myall_leaves <- paste("subgroup", seq(1,40), sep="_")
myconnect <- rbind( 
  data.frame( from=sample(myall_leaves, 40, replace=T) , to=sample(myall_leaves, 40, replace=T)), 
  data.frame( from=sample(head(myall_leaves), 20, replace=T) , to=sample( tail(myall_leaves), 20, replace=T)), 
  data.frame( from=sample(myall_leaves[21:24], 10, replace=T) , to=sample( myall_leaves[25:30], 10, replace=T)))
myconnect$value <- runif(nrow(myconnect))
```


 
## create a vertices data.frame. One line per object of our hierarchy
```{r}
myvertices  <-  data.frame(
  name = unique(c(as.character(myedges$from), as.character(myedges$to))) , 
  value = runif(49)
) 
```


## Let's add a column with the group of each name. It will be useful later to color points
The group column specifies where the vertex come from, or a closes higher level. For instance, origin will have NA, while group1 and group2 will have origin as a group (as closest next level above).

```{r}
myvertices$group  <-  myedges$from[ match( myvertices$name, myedges$to ) ]
```


# Creating the labels
Next step: computing the label features that will be displayed all around the circle, next to the nodes:

angle → vertical on top and botton, horizontal on the side, and so on.
flip it → labels on the left hand side must be 180° flipped to be readable
alignment → if labels are flipped, they must be right aligned
Those information are computed and added to the vertices data frame.

#Let's add information concerning the label we are going to add: angle, horizontal adjustement and potential flip
#calculate the ANGLE of the labels

leaves is everything what is not sectors, not the main level. Here origin and sectors are not leaves.

I think i can create in Excel the vertices dataframe, with all the same features. Then just import it here.
```{r}
myvertices$id <- NA
myleaves <- which(is.na( match(myvertices$name, myedges$from) ))
mynleaves <- length(myleaves)
myvertices$id[ myleaves ] <- seq(1:mynleaves)
myvertices$angle <- 90 - 360 * myvertices$id / mynleaves
```


 
## calculate the alignment of labels: right or left
 If I am on the left part of the plot, my labels have currently an angle < -90
```{r}
myvertices$hjust <- ifelse( myvertices$angle < -90, 1, 0)
```


## flip angle BY to make them readable
```{r}
myvertices$angle <- ifelse(myvertices$angle < -90, myvertices$angle+180, myvertices$angle)

```

## Trying to make the angles for the lables upright 
```{r}
myvertices$angle2 <-myvertices$angle + 90
```



## No color
```{r  fig.width=7, fig.height=7}
# Create a graph object
mygraph2 <- igraph::graph_from_data_frame( myedges, vertices=myvertices )
 
# The connection object must refer to the ids of the leaves:
from  <-  match( myconnect$from, myvertices$name)
to  <-  match( myconnect$to, myvertices$name)
 
# Basic usual argument
ggraph(mygraph2, layout = 'dendrogram', circular = TRUE) + 
  geom_node_point(aes(filter = leaf, x = x*1.05, y=y*1.05)) +
  geom_conn_bundle(data = get_con(from = from, to = to), alpha=0.2, colour="skyblue", width=0.9) +
  geom_node_text(aes(x = x*1.1, y=y*1.1, filter = leaf, label=name, angle = angle, hjust=hjust), size=1.5, alpha=1) +
  theme_void() +
  theme(
    legend.position="none",
    plot.margin=unit(c(0,0,0,0),"cm"),
  ) +
  expand_limits(x = c(-1.2, 1.2), y = c(-1.2, 1.2))

```

?? How to change the angle to 180 for each?


## With color
```{r fig.width=7, fig.height=7}
ggraph(mygraph2, layout = "dendrogram", circular = TRUE) +
  geom_conn_bundle(data = get_con(from = from, to = to), alpha = 0.2, width = 0.9, aes(colour = ..index..)) +
  scale_edge_colour_distiller(palette = "RdPu") +

  geom_node_text(aes(x = x * 1.15, y = y * 1.15, filter = leaf, label = name, angle = angle, hjust = hjust, colour = group), size = 2, alpha = 1) +

  geom_node_point(aes(filter = leaf, x = x * 1.07, y = y * 1.07, colour = group, size = value, alpha = 0.2)) +
  scale_colour_manual(values = rep(brewer.pal(9, "Paired"), 30)) +
  scale_size_continuous(range = c(0.1, 10)) +

  theme_void() +
  theme(
    legend.position = "none",
    plot.margin = unit(c(0, 0, 0, 0), "cm"),
  ) +
  expand_limits(x = c(-1.3, 1.3), y = c(-1.3, 1.3))
```



I will have to create two dataframes by hand: myconnect and myvertices in Excel. See carefully, what each one represents, it is not difficult after all.


# Benjamin's angle calculations

```{r }
#a little cryptic but I hope it can help:

geom_node_text(aes(x = x*0.95, y=y*0.95, filter = !leaf,
angle = pmax((acos(x/(((x^2)+(y^2))^0.5))*sign(y)*180/pi-((sign(x)==-1)*(sign(y)==1)*180)+((sign(x)==-1)*(sign(y)==-1)*180)),-180,na.rm=TRUE),
hjust = (sign(x)==1),
label=name),
size=2.5,
alpha=1,
color = 'darkslategrey',
repel=FALSE) + #legends
```

```{r fig.width=5, fig.height=5}
ggraph(mygraph2, layout = 'dendrogram', circular = TRUE) + 
  geom_conn_bundle(data = get_con(from = from, to = to), alpha=0.2, width=0.9, aes(colour=..index..)) +
  scale_edge_colour_distiller(palette = "RdPu") +
  
  geom_node_text(aes(x = x*1.3, y=y*1.3, filter = leaf, label=name, 
                     angle = pmax((acos(x/(((x^2)+(y^2))^0.5))*sign(y)*180/pi-((sign(x)==-1)*(sign(y)==1)*180)+((sign(x)==-1)*(sign(y)==-1)*180)),-180,na.rm=TRUE), 
                     hjust=(sign(x)==1), 
                     colour=group), size=3, alpha=1) +
  
  geom_node_point(aes(filter = leaf, x = x*0.9, y=y*0.9, colour=group, size=value, alpha=0.2)) +
  scale_colour_manual(values= rep( brewer.pal(9,"Paired") , 30)) +
  scale_size_continuous( range = c(0.1,10) ) +
  
  theme_void() +
  theme(
    legend.position="none",
    plot.margin=unit(c(0,0,0,0),"cm"),
  ) +
  expand_limits(x = c(-1.3, 1.3), y = c(-1.3, 1.3))
```

